---
title: "25-26_doe_data"
format: html
---

## Import package and trying to connect snowflake

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(purrr)
library(stringdist)
library(openxlsx)
library(stringr)
#The following package is for connecting to Snowflake.
library(DBI)
library(odbc)
library(openssl)  # To read private key
```

## Connect to Snowflake

```{r}
# Connect to Snowflake using key-pair authentication

con <- dbConnect(odbc::odbc(),
                 Driver = "SnowflakeDSIIDriver", #from UFIT
                 Server = "ufl-datahub.snowflakecomputing.com", #from LC Data team
                 UID = "xu.rui@ufl.edu", #from UFIT
                 authenticator = "externalbrowser", #from UFIT
                 Warehouse     = "REPORTING", #from LC Data team
                 Database      = "PUB_LAKE_LASTINGER" #can specify specific database needing to access
                #Schema        = "ADD" #can specify specific Schema - aka folders - needing to access
)

# Look in Connections - top right RStudio - for Database/Schema list
```

## Import data in snowflake

```{r}
staff <- tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_STAFF"))
```

```{r}
staff_suffix <- tbl(con,dbplyr::in_schema("RAW_DOE","02_13_2026_PERA_4472B_STAFF_CERTIFICATION_DATA_2025_26_FS2"))
```

```{r}
teacher_course <- tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_TEACHER_COURSE"))
```

```{r}
student_demo <-  tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_STUDENT_DEMO"))
```

import data in local data

```{r}

raw_Literacy_Tx_UserforDoe <- read_csv("raw_Literacy_Tx_UserforDoe.csv")
```

```{r}
msid_all_school <- read_excel("msid_all_school.xlsx")
```

## section 1: combine two staff data sets

## explore

```{r}
staff %>%
  count(TEACHER_ID) %>%
  filter(n > 1)
```

```{r}
qc_teacher_id <- staff %>%
  summarise(
    total_rows = n(),
    unique_teacher = n_distinct(TEACHER_ID),
    missing_teacher = sum(ifelse(is.na(TEACHER_ID), 1L, 0L), na.rm = TRUE)#note: The is.na function cannot be recognized in Snowflake. Therefore, we must switch to a method that is compatible with sql.
  ) %>%
  collect()

qc_teacher_id


```

```{r}
qc_teacher_id_1 <- staff_suffix %>%
  summarise(
    total_rows = n(),
    unique_teacher = n_distinct(TEACHER_ID),
    missing_teacher = sum(ifelse(is.na(TEACHER_ID), 1L, 0L), na.rm = TRUE)#same issue
  ) %>%
  collect()

qc_teacher_id_1

```

Filter out duplicate data\
function:

```{r}


# =========================================================
# Function: split unique TEACHER_ID rows and duplicated rows
# 功能：
# 1. 根据 TEACHER_ID 统计重复情况
# 2. 保留唯一 TEACHER_ID 的数据
# 3. 把重复 TEACHER_ID 的数据单独保存
# 4. 给数据加上来源标签 source
# =========================================================
split_teacher_data <- function(df, source_name) {
  
  # 第一步：计算每个 TEACHER_ID 出现的次数
  teacher_count <- df %>%
    count(TEACHER_ID, name = "teacher_id_n")
  
  # 第二步：把次数合并回原始数据
  df_with_count <- df %>%
    left_join(teacher_count, by = "TEACHER_ID")
  
  # 第三步：保留 TEACHER_ID 只出现 1 次的行
  df_unique <- df_with_count %>%
    filter(teacher_id_n == 1) %>%
    mutate(source = source_name)
  
  # 第四步：把重复的 TEACHER_ID 行单独保存
  df_duplicate <- df_with_count %>%
    filter(teacher_id_n > 1) %>%
    mutate(source = source_name)
  
  # 第五步：重复 TEACHER_ID 的 summary（每个ID重复几次）
  dup_summary <- teacher_count %>%
    filter(teacher_id_n > 1) %>%
    arrange(desc(teacher_id_n))
  
  # 返回一个 list，方便后续调用
  return(list(
    unique_data = df_unique,
    duplicate_data = df_duplicate,
    duplicate_summary = dup_summary
  ))
}
```

```{r}
# 对第一个数据库处理
staff_2526_result <- split_teacher_data(
  df = staff,
  source_name = "staff"
)

# 对第二个数据库处理
staff_july_result <- split_teacher_data(
  df = staff_suffix,
  source_name = "staff_suffix"
)
```

```{r}
# 第一个数据库：唯一 / 重复 / 重复summary
staff_2526_unique <- staff_2526_result$unique_data
staff_2526_duplicate <- staff_2526_result$duplicate_data
staff_2526_dup_summary <- staff_2526_result$duplicate_summary

# 第二个数据库：唯一 / 重复 / 重复summary
staff_july_unique <- staff_july_result$unique_data
staff_july_duplicate <- staff_july_result$duplicate_data
staff_july_dup_summary <- staff_july_result$duplicate_summary
```

export duplication

```{r}
write.csv(staff_2526_duplicate, "staff_2526_duplicate.csv", row.names = FALSE)
write.csv(staff_july_duplicate, "staff_suffix.csv", row.names = FALSE)
```

export unique teacher\

```{r}
write.csv(staff_2526_unique,"staff_2526_uqniue.csv", row.names = FALSE)
write.csv(staff_july_unique,"staff_suffix_unique.csv", row.names = FALSE)
```

function add suffix

```{r}
# Function: add suffix to all variables except key columns
# 功能：
# 给所有列加后缀，但保留 key（默认 TEACHER_ID 不变）
# =========================================================
add_suffix_except <- function(df, suffix, keep = "TEACHER_ID") {
  
  df_renamed <- df %>%
    rename_with(
      .fn = ~ ifelse(.x %in% keep, .x, paste0(.x, suffix))
    )
  
  return(df_renamed)
}
```

```{r}
staff_2526_unique_sfx <- add_suffix_except(
  df = staff_2526_unique,
  suffix = "_2526"
)
```

## unique data merge

```{r}
staff_merged <- staff_july_unique %>%
  full_join(
    staff_2526_unique_sfx,
    by = "TEACHER_ID"
  )
```

## qc

```{r}
staff_merged %>%
  summarise(
    total_rows = n(),
    unique_teacher = n_distinct(TEACHER_ID)
  )
```

```{r}

staff_merged_qc <- staff_merged %>%
  mutate(
    source_flag = case_when(
      !is.na(source) & !is.na(source_2526) ~ "both",
      !is.na(source) & is.na(source_2526) ~ "only_july",
      is.na(source) & !is.na(source_2526) ~ "only_2526",
      TRUE ~ NA_character_
    )
  )

```

## For duplication data merge

Check the combination of district and teacher id

```{r}
staff_july_duplicate_check <- staff_july_duplicate %>%
  count(TEACHER_ID, DISTRICT_SERVICE, name = "n") %>%
  filter(n > 1) %>%
  arrange(desc(n))

staff_july_duplicate_check
```

```{r}
staff_2526_duplicate_check <- staff_2526_duplicate %>%
  count(TEACHER_ID, DISTRICT_SERVICE, name = "n") %>%
  filter(n > 1) %>%
  arrange(desc(n))

staff_2526_duplicate_check
```

```{r}
staff_july_duplicate %>%
  summarise(
    total_rows = n(),
    distinct_teacher = n_distinct(TEACHER_ID),
    distinct_teacher_district = n_distinct(paste(TEACHER_ID, DISTRICT_SERVICE, sep = "__"))
  )
```

```{r}
staff_2526_duplicate %>%
  summarise(
    total_rows = n(),
    distinct_teacher = n_distinct(TEACHER_ID),
    distinct_teacher_district = n_distinct(paste(TEACHER_ID, DISTRICT_SERVICE, sep = "__"))
  )
```

```{r}
nrow(staff_july_duplicate_check)
nrow(staff_2526_duplicate_check)
```

```{r}
staff_2526_duplicate_sfx <- add_suffix_except(
  df = staff_2526_duplicate,
  suffix = "_2526",
  keep = c("TEACHER_ID", "DISTRICT_SERVICE")
)
```

```{r}
staff_duplicate_merged <- staff_july_duplicate %>%
  full_join(
    staff_2526_duplicate_sfx,
    by = c("TEACHER_ID", "DISTRICT_SERVICE")
  )
```

```{r}
staff_duplicate_merged <- staff_duplicate_merged %>%
  mutate(
    source_flag = case_when(
      !is.na(source) & !is.na(source_2526) ~ "both",
      !is.na(source) & is.na(source_2526) ~ "only_july",
      is.na(source) & !is.na(source_2526) ~ "only_2526",
      TRUE ~ NA_character_
    )
  )
```

```{r}
staff_duplicate_merged %>%
  count(source_flag)
```

```{r}
setdiff(colnames(staff_merged), colnames(staff_duplicate_merged))
setdiff(colnames(staff_duplicate_merged), colnames(staff_merged))
```

```{r}
staff_duplicate_merged <- staff_duplicate_merged %>%
  mutate(DISTRICT_SERVICE_2526 = DISTRICT_SERVICE)
```

```{r}
#note: The bind_row function cannot be used directly on Snowflake. Therefore, we need to replace it with a compatible Snowflake function called union_all.
staff_final_merged <- union_all(
  staff_merged,
  staff_duplicate_merged
) %>% 
  collect()
```

```{r}
staff_final_merged %>% count(teacher_id_n_2526)
```

```{r}
write.csv(staff_final_merged,
      "survey2_2526_twodatasets_merge.csv",
row.names = FALSE)
```

```{r}
staff_final_merged <- read_csv("survey2_2526_twodatasets_merge.csv")
```

## section 2 merge new internal data the first match is using certificate number

```{r}
#First, we have filtered out the treatment datasets for the three programs in the 2025-26 academic year.
df_literacy <- raw_Literacy_Tx_UserforDoe %>%
  filter(PL_2526_Tx == 1|
         LM_2526_Tx == 1|
         MC_2526_Tx == 1 ) %>%
  mutate(
    CERT_1_clean = str_to_upper(
      str_replace_all(as.character(CERT_1), "[^A-Za-z0-9]", "")
    ),
    CERT_2_clean = str_to_upper(
      str_replace_all(as.character(CERT_2), "[^A-Za-z0-9]", "")
    ),
    CERT_3_clean = str_to_upper(
      str_replace_all(as.character(CERT_3), "[^A-Za-z0-9]", "")
    )
  )
#To prevent issues during data merging, we're deleting entries from the staff dataset that don't have a certificate number.
df_cert <- staff_final_merged %>%
  mutate(
    Cert_clean = str_to_upper(
      str_replace_all(
        as.character(CERTIFICATE_FILE_NUMBER),
        "[^A-Za-z0-9]",
        ""
      )
    )
  ) %>%
  filter(!is.na(Cert_clean), Cert_clean != "") %>%
  distinct(Cert_clean, .keep_all = TRUE)
```

```{r}
df_literacy %>%
  count(
    PL_2526_Tx,
    LM_2526_Tx,
    MC_2526_Tx,
    sort = TRUE
  )
```

```{r}
df_cert %>%
  count(Cert_clean, sort = TRUE) %>%
  filter(n > 1)
```

```{r}
df_literacy %>%
  count(CERT_1_clean, sort = TRUE) %>%
  filter(n > 1) %>% 
  print(Inf)
```

## matching part 1  : certificate 

```{r}
cert_match_1 <- df_literacy %>%
  left_join(
    df_cert,
    by = c("CERT_1_clean" = "Cert_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(CERTIFICATE_FILE_NUMBER),
      "matched_by_CERT_1",
      NA_character_
    )
  )

```

```{r}
cert_no_match_1 <- cert_match_1 %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_literacy))
```

```{r}
cert_match_2 <- cert_no_match_1 %>%
  left_join(
    df_cert,
    by = c("CERT_2_clean" = "Cert_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(CERTIFICATE_FILE_NUMBER),
      "matched_by_CERT_2",
      NA_character_
    )
  )

# 找出第二轮仍然没有匹配上的原始 Literacy 行
cert_no_match_2 <- cert_match_2 %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_literacy))
```

```{r}
# Step 5: 第三轮，用 CERT_3 匹配第二轮失败的个案
cert_match_3 <- cert_no_match_2 %>%
  left_join(
    df_cert,
    by = c("CERT_3_clean" = "Cert_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(CERTIFICATE_FILE_NUMBER),
      "matched_by_CERT_3",
      NA_character_
    )
  )
```

```{r}
cert_final_match <- bind_rows(
  cert_match_1 %>% filter(!is.na(CERTIFICATE_FILE_NUMBER)),
  cert_match_2 %>% filter(!is.na(CERTIFICATE_FILE_NUMBER)),
  cert_match_3
)
```

```{r}
cert_no_match <- cert_final_match %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_literacy))
```

```{r}
cert_final_match %>%
  count(match_type)
```

```{r}
cert_no_match %>%
  count(CERT_1_clean)
```

##### 

After updating the treatment dataset, it now contains a total of 7,721 observations.

Even after merging by certificate number, there are still 3,258 observations that could not be matched. For the next step, I will use email addresses to perform the subsequent merge.

## certificate_2526

```{r}
cert_no_match
```

## this section: final_no_match I'll use the email variable to perform the merge again.

```{r}
# Step 1: 先清理 final_no_match 里的 5 个 email 变量
df_email_no_match <- final_no_match %>%
  mutate(
    literacy_email_row_id = row_number(),
    
    username_clean = str_to_lower(str_trim(as.character(username))),
    username_2_clean = str_to_lower(str_trim(as.character(username_2))),
    WORKEMAIL_1_clean = str_to_lower(str_trim(as.character(WORKEMAIL_1))),
    WORKEMAIL_2_clean = str_to_lower(str_trim(as.character(WORKEMAIL_2))),
    WORKEMAIL_3_clean = str_to_lower(str_trim(as.character(WORKEMAIL_3)))
  )
```

```{r}
# Step 2: 清理 df_cert 里的 EMAIL_ADDRESS_2526
df_cert_email <- df_cert %>%
  mutate(
    EMAIL_ADDRESS_2526_clean = str_to_lower(
      str_trim(as.character(EMAIL_ADDRESS_2526))
    )
  ) %>%
  filter(
    !is.na(EMAIL_ADDRESS_2526_clean),
    EMAIL_ADDRESS_2526_clean != ""
  ) %>%
  distinct(EMAIL_ADDRESS_2526_clean, .keep_all = TRUE)
```

```{r}
email_match_1 <- df_email_no_match %>%
  left_join(
    df_cert_email,
    by = c("username_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_username",
      NA_character_
    )
  )

email_no_match_1 <- email_match_1 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_2 <- email_no_match_1 %>%
  left_join(
    df_cert_email,
    by = c("username_2_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_username_2",
      NA_character_
    )
  )

email_no_match_2 <- email_match_2 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_3 <- email_no_match_2 %>%
  left_join(
    df_cert_email,
    by = c("WORKEMAIL_1_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_WORKEMAIL_1",
      NA_character_
    )
  )

email_no_match_3 <- email_match_3 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_4 <- email_no_match_3 %>%
  left_join(
    df_cert_email,
    by = c("WORKEMAIL_2_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_WORKEMAIL_2",
      NA_character_
    )
  )

email_no_match_4 <- email_match_4 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_5 <- email_no_match_4 %>%
  left_join(
    df_cert_email,
    by = c("WORKEMAIL_3_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_WORKEMAIL_3",
      NA_character_
    )
  )
```

```{r}
email_final_match <- bind_rows(
  email_match_1 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_2 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_3 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_4 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_5
)
```

```{r}
email_final_no_match <- email_final_match %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_final_match %>%
  count(match_type) %>% 
  print(n = Inf, width = Inf)

```

There are still about 300 records here that cannot be merged. Given the situation, I think merging based solely on first and last name is too risky. I am considering whether we should include the MSID to create a new combination of first name, last name, school, and district to complete the merging.

```{r}
cert_success <- final_match %>%
  filter(!is.na(CERTIFICATE_FILE_NUMBER)) %>%
  mutate(match_stage = "certificate")
```

```{r}
email_success <- email_final_match %>%
  filter(!is.na(EMAIL_ADDRESS_2526)) %>%
  mutate(match_stage = "email")
```

```{r}
all_success <- bind_rows(
  cert_success,
  email_success
)
```

```{r}
all_failed <- email_final_match %>%
  filter(is.na(EMAIL_ADDRESS_2526))
```

```{r}
all_success %>%
  count(match_type, sort = TRUE) %>% 
  print(n = Inf, width = Inf)
```

```{r}
write.csv(all_success,"pl_2526_txwith_staff_inf.csv",row.names = FALSE)
```

this is the tempory attempt to continue merging data

```{r}
names(msid_all_school)
```

```{r}
all_failed %>% 
  count(COUNTY_1)
```

```{r}
msid <- msid_all_school %>% 
  add_suffix_except(
    suffix = "_msid"
  )
```

```{r}
msid_district <- msid %>%
  mutate(
    district_name_clean = str_to_upper(str_trim(DISTRICT_NAME_msid))
  ) %>%
  filter(
    !is.na(district_name_clean),
    district_name_clean != ""
  ) %>%
  distinct(
    district_name_clean,
    DISTRICT_msid,
    .keep_all = TRUE
  )

```

```{r}
msid_district <- msid_district %>% 
  select(DISTRICT_msid,district_name_clean)
```

```{r}
# Step 2: 清理 all_failed 里的 COUNTY_1/2/3
all_failed_county <- all_failed %>%
  mutate(
    failed_row_id = row_number(),
    
    COUNTY_1_clean = str_to_upper(str_trim(as.character(COUNTY_1))),
    COUNTY_2_clean = str_to_upper(str_trim(as.character(COUNTY_2))),
    COUNTY_3_clean = str_to_upper(str_trim(as.character(COUNTY_3)))
  )
```

```{r}
county_match_1 <- all_failed_county %>%
  left_join(
    msid_district,
    by = c("COUNTY_1_clean" = "district_name_clean")
  ) %>%
  mutate(
    county_match_type = if_else(
      !is.na(DISTRICT_msid),
      "matched_by_COUNTY_1",
      NA_character_
    )
  )

county_no_match_1 <- county_match_1 %>%
  filter(is.na(DISTRICT_msid)) %>%
  select(names(all_failed_county))
```

```{r}
county_match_2 <- county_no_match_1 %>%
  left_join(
    msid_district,
    by = c("COUNTY_2_clean" = "district_name_clean")
  ) %>%
  mutate(
    county_match_type = if_else(
      !is.na(DISTRICT_msid),
      "matched_by_COUNTY_2",
      NA_character_
    )
  )

county_no_match_2 <- county_match_2 %>%
  filter(is.na(DISTRICT_msid)) %>%
  select(names(all_failed_county))
```

```{r}
county_match_3 <- county_no_match_2 %>%
  left_join(
    msid_district,
    by = c("COUNTY_3_clean" = "district_name_clean")
  ) %>%
  mutate(
    county_match_type = if_else(
      !is.na(DISTRICT_msid),
      "matched_by_COUNTY_3",
      NA_character_
    )
  )
```

```{r}
all_failed_with_district <- bind_rows(
  county_match_1 %>% filter(!is.na(DISTRICT_msid)),
  county_match_2 %>% filter(!is.na(DISTRICT_msid)),
  county_match_3
)
```

```{r}
all_failed_with_district %>%
  count(county_match_type) %>%
  print(n = Inf)
```

```{r}
names(all_failed_with_district)
```

```{r}
names(staff_final_merged)
```

```{r}
# Step 1: 清理 all_failed_with_district
failed_name_district <- all_failed_with_district %>%
  mutate(
    NAME_key_first_clean = str_to_upper(str_trim(as.character(NAME_key_first))),
    NAME_key_last_clean  = str_to_upper(str_trim(as.character(NAME_key_last))),
    DISTRICT_msid_clean  = str_pad(
      as.character(DISTRICT_msid),
      width = 2,
      side = "left",
      pad = "0"
    )
  )
```

```{r}
# Step 2: 清理 staff_final_merged
staff_name_district <- staff_final_merged %>%
  mutate(
    FIRST_NAME_clean = str_to_upper(str_trim(as.character(FIRST_NAME))),
    LAST_NAME_clean  = str_to_upper(str_trim(as.character(LAST_NAME))),
    DISTRICT_SERVICE_2526_clean = str_pad(
      as.character(DISTRICT_SERVICE_2526),
      width = 2,
      side = "left",
      pad = "0"
    )
  ) %>%
  filter(
    !is.na(FIRST_NAME_clean),
    FIRST_NAME_clean != "",
    !is.na(LAST_NAME_clean),
    LAST_NAME_clean != "",
    !is.na(DISTRICT_SERVICE_2526_clean),
    DISTRICT_SERVICE_2526_clean != ""
  )
```

```{r}
staff_name_district %>%
  count(FIRST_NAME_clean, LAST_NAME_clean, DISTRICT_SERVICE_2526_clean, sort = TRUE) %>%
  filter(n > 1)
```

```{r}
name_district_match <- failed_name_district %>%
  left_join(
    staff_name_district,
    by = c(
      "NAME_key_first_clean" = "FIRST_NAME_clean",
      "NAME_key_last_clean"  = "LAST_NAME_clean",
      "DISTRICT_msid_clean"  = "DISTRICT_SERVICE_2526_clean"
    )
  ) 
```

```{r}

name_district_no_match <- name_district_match %>%
  filter(is.na(teacher_id_n.y))

```

## Section 3 combing student data

```{r}
#Load TeacherCourse locally.
teacher_course_local <- teacher_course %>% collect()
```

```{r}
teacher_course_local <- teacher_course_local %>% 
  add_suffix_except(
  suffix = "_tc",
  keep = "TEACHER_ID"
  )
```

```{r}
treatment_data <- all_success %>% 
  left_join(teacher_course_local,
            by=c("TEACHER_ID",
            "DISTRICT_SERVICE"="DISTRICT_INST_tc"))
```

```{r}
treatment_data %>%
  count(TEACHER_ID, DISTRICT_SERVICE) %>%
  filter(n > 1)
```

```{r}
treatment_data %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  summarise(n = n())
```

```{r}
treatment_no_studentinfo <- treatment_data %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc))
```

```{r}
missing_fle <- treatment_data %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  select(TEACHER_ID) %>%
  distinct()
```

```{r}
teacher_course_check <- teacher_course_local %>%
  filter(TEACHER_ID %in% missing_fle$TEACHER_ID)
```

```{r}
write.csv(treatment_no_studentinfo,
        "teachers_without_students_info.csv",
       row.names = FALSE)
```

The inspection is complete. I have confirmed that these teacher IDs are not in the teacher Courses data. need to double check and discuss how to handle this issue.

next , I will combine student demo and deduplication

```{r}
treatment_data <- treatment_data %>% 
  mutate(
    DISTRICT_tc = as.numeric(as.character(DISTRICT_tc)),
    SCHOOL_tc   = as.numeric(as.character(SCHOOL_tc)),
    SchoolID_tc = DISTRICT_tc * 10000 + SCHOOL_tc,
    UniqueID    = paste0(FLEID_tc, SchoolID_tc)
  )
```

```{r}
student_demo_local <- student_demo %>% collect()
```

```{r}
student_demo_local <- student_demo_local %>% 
  add_suffix_except(
  suffix = "_demo",
  keep = "FLEID"
  )
```

```{r}
student_demo_local <- student_demo_local %>% 
  mutate(
    DISTRICT_demo = as.numeric(as.character(DISTRICT_demo)),
    SCHOOL_demo   = as.numeric(as.character(SCHOOL_demo)),
    SchoolID_demo = DISTRICT_demo * 10000 + SCHOOL_demo,
    UniqueID    = paste0(FLEID, SchoolID_demo)
  )
```

```{r}
dup_in_demo <- student_demo_local %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  pull(UniqueID)
```

deduplication strategy 1 :deduplication in student_demo drop district_inst ==71

```{r}
student_demo_clean <- student_demo_local %>%
  group_by(UniqueID) %>%
  filter(!(n() > 1 & DISTRICT_INST_demo == 71)) %>%
  ungroup()
```

```{r}
nrow(student_demo_local)
nrow(student_demo_clean)
```

```{r}
student_demo_clean %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  arrange(desc(n))# need to manually check
```

```{r}
treatment_data_final <- treatment_data %>% 
  left_join(student_demo_clean,
            by="UniqueID")
```

```{r}
 treatment_data_final %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  arrange(desc(n))
```

```{r}
treatment_data_final %>% 
  filter(UniqueID =="FL000007593932180301")
```

```{r}
write.csv(treatment_data_final,
          "main_treatment_data_initial.csv",
          row.names = FALSE)
```

```{r}
dbDisconnect(con)
```
