---
title: "25_26new_internal_data_v2"
format: html
---

## Import package and trying to connect snowflake

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(purrr)
library(stringdist)
library(openxlsx)
library(stringr)
#The following package is for connecting to Snowflake.
library(DBI)
library(odbc)
library(openssl)  # To read private key
```

## Connect to Snowflake

```{r}
# Connect to Snowflake using key-pair authentication

con <- dbConnect(odbc::odbc(),
                 Driver = "SnowflakeDSIIDriver", #from UFIT
                 Server = "ufl-datahub.snowflakecomputing.com", #from LC Data team
                 UID = "xu.rui@ufl.edu", #from UFIT
                 authenticator = "externalbrowser", #from UFIT
                 Warehouse     = "REPORTING", #from LC Data team
                 Database      = "PUB_LAKE_LASTINGER" #can specify specific database needing to access
                #Schema        = "ADD" #can specify specific Schema - aka folders - needing to access
)

# Look in Connections - top right RStudio - for Database/Schema list
```

## Import data in snowflake

```{r}
staff <- tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_STAFF"))
```

```{r}
staff_suffix <- tbl(con,dbplyr::in_schema("RAW_DOE","02_13_2026_PERA_4472B_STAFF_CERTIFICATION_DATA_2025_26_FS2"))
```

```{r}
teacher_course <- tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_TEACHER_COURSE"))
```

```{r}
student_demo <-  tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_STUDENT_DEMO"))
```

import data in local data

```{r}

raw_Literacy_Tx_UserforDoe <- read_csv(
  "raw_Literacy_Tx_UserforDoe.csv",
  locale = locale(encoding = "Latin1")
)
```

```{r}
msid_all_school <- read_excel("msid_all_school.xlsx")
```

section1 is deleted due to avoiding memory issue

```{r}
staff_final_merged <- read_csv("survey2_2526_twodatasets_merge.csv")
```

## section 2 merge new internal data the first match is using certificate number

```{r}
#First, we have filtered out the treatment datasets for the three programs in the 2025-26 academic year.
df_literacy <- raw_Literacy_Tx_UserforDoe %>%
  filter(PL_2526_Tx == 1|
         LM_2526_Tx == 1|
         MC_2526_Tx == 1 ) %>%
  mutate(
    CERT_1_clean = str_to_upper(
      str_replace_all(as.character(CERT_1), "[^A-Za-z0-9]", "")
    ),
    CERT_2_clean = str_to_upper(
      str_replace_all(as.character(CERT_2), "[^A-Za-z0-9]", "")
    ),
    CERT_3_clean = str_to_upper(
      str_replace_all(as.character(CERT_3), "[^A-Za-z0-9]", "")
    )
  )
#To prevent issues during data merging, we're deleting entries from the staff dataset that don't have a certificate number.
df_cert <- staff_final_merged %>%
  mutate(
    Cert_clean = str_to_upper(
      str_replace_all(
        as.character(CERTIFICATE_FILE_NUMBER),
        "[^A-Za-z0-9]",
        ""
      )
    )
  ) %>%
  filter(!is.na(Cert_clean), Cert_clean != "") %>%
  distinct(Cert_clean, .keep_all = TRUE)
```

```{r}
df_literacy %>%
  count(
    PL_2526_Tx,
    LM_2526_Tx,
    MC_2526_Tx,
    sort = TRUE
  )
```

```{r}
df_cert %>%
  count(Cert_clean, sort = TRUE) %>%
  filter(n > 1)
```

```{r}
df_literacy %>%
  count(CERT_1_clean, sort = TRUE) %>%
  filter(n > 1) %>% 
  print(Inf)
```

## matching part 1 : certificate vs certificate_file_number

```{r}
cert_match_1 <- df_literacy %>%
  left_join(
    df_cert,
    by = c("CERT_1_clean" = "Cert_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(CERTIFICATE_FILE_NUMBER),
      "matched_by_CERT_1",
      NA_character_
    )
  )

```

```{r}
cert_no_match_1 <- cert_match_1 %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_literacy))
```

```{r}
cert_match_2 <- cert_no_match_1 %>%
  left_join(
    df_cert,
    by = c("CERT_2_clean" = "Cert_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(CERTIFICATE_FILE_NUMBER),
      "matched_by_CERT_2",
      NA_character_
    )
  )

# 找出第二轮仍然没有匹配上的原始 Literacy 行
cert_no_match_2 <- cert_match_2 %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_literacy))
```

```{r}
# Step 5: 第三轮，用 CERT_3 匹配第二轮失败的个案
cert_match_3 <- cert_no_match_2 %>%
  left_join(
    df_cert,
    by = c("CERT_3_clean" = "Cert_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(CERTIFICATE_FILE_NUMBER),
      "matched_by_CERT_3",
      NA_character_
    )
  )
```

```{r}
cert_final_match <- bind_rows(
  cert_match_1 %>% filter(!is.na(CERTIFICATE_FILE_NUMBER)),
  cert_match_2 %>% filter(!is.na(CERTIFICATE_FILE_NUMBER)),
  cert_match_3
)
```

```{r}
cert_no_match <- cert_final_match %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_literacy))
```

```{r}
cert_final_match %>%
  count(match_type)
```

```{r}
cert_no_match %>%
  count(CERT_1_clean)
```

##### 

After updating the treatment dataset, it now contains a total of 7,721 observations.

Even after merging by certificate number, there are still 3,258 observations that could not be matched. For the next step, I will use certificate vs certificate_2526 to perform the subsequent merge.

## certificate_2526(this variable is from staff_survey_2)

```{r}
df_cert_2526 <- staff_final_merged %>%
  mutate(
    Cert_clean_2526 = str_to_upper(
      str_replace_all(
        as.character(CERTIFICATE_2526),
        "[^A-Za-z0-9]",
        ""
      )
    )
  ) %>%
  filter(!is.na(Cert_clean_2526), Cert_clean_2526 != "") %>%
  distinct(Cert_clean_2526, .keep_all = TRUE)
```

```{r}
# Step 1: 确保 no match 数据里有干净的 CERT_1/2/3
cert_no_match_2526 <- cert_no_match %>%
  mutate(
    no_match_row_id = row_number(),
    CERT_1_clean = str_to_upper(str_replace_all(as.character(CERT_1), "[^A-Za-z0-9]", "")),
    CERT_2_clean = str_to_upper(str_replace_all(as.character(CERT_2), "[^A-Za-z0-9]", "")),
    CERT_3_clean = str_to_upper(str_replace_all(as.character(CERT_3), "[^A-Za-z0-9]", ""))
  )
```

```{r}
# Step 2: 用 CERT_1 匹配 df_cert_2526
cert2526_match_1 <- cert_no_match_2526 %>%
  left_join(
    df_cert_2526,
    by = c("CERT_1_clean" = "Cert_clean_2526")
  ) %>%
  mutate(
    match_type_2526 = if_else(
      !is.na(CERTIFICATE_2526),
      "matched_2526_by_CERT_1",
      NA_character_
    )
  )

cert2526_no_match_1 <- cert2526_match_1 %>%
  filter(is.na(CERTIFICATE_2526)) %>%
  select(names(cert_no_match_2526))
```

```{r}
# Step 3: CERT_1 没匹配上的，再用 CERT_2
cert2526_match_2 <- cert2526_no_match_1 %>%
  left_join(
    df_cert_2526,
    by = c("CERT_2_clean" = "Cert_clean_2526")
  ) %>%
  mutate(
    match_type_2526 = if_else(
      !is.na(CERTIFICATE_2526),
      "matched_2526_by_CERT_2",
      NA_character_
    )
  )

cert2526_no_match_2 <- cert2526_match_2 %>%
  filter(is.na(CERTIFICATE_2526)) %>%
  select(names(cert_no_match_2526))
```

```{r}
# Step 4: CERT_2 也没匹配上的，再用 CERT_3
cert2526_match_3 <- cert2526_no_match_2 %>%
  left_join(
    df_cert_2526,
    by = c("CERT_3_clean" = "Cert_clean_2526")
  ) %>%
  mutate(
    match_type_2526 = if_else(
      !is.na(CERTIFICATE_2526),
      "matched_2526_by_CERT_3",
      NA_character_
    )
  )
```

```{r}
cert2526_final_match <- bind_rows(
  cert2526_match_1 %>% filter(!is.na(CERTIFICATE_2526)),
  cert2526_match_2 %>% filter(!is.na(CERTIFICATE_2526)),
  cert2526_match_3
)
```

```{r}
cert2526_final_match %>%
  count(match_type_2526) %>%
  print(n = Inf, width = Inf)
```

```{r}
cert2526_final_no_match <- cert2526_final_match %>%
  filter(is.na(CERTIFICATE_2526)) %>%
  select(names(cert_no_match_2526))
```

combine togeteher

```{r}
# 1. 原来 certificate 成功匹配的部分
cert_success_old <- cert_final_match %>%
  filter(!is.na(CERTIFICATE_FILE_NUMBER)) %>%
  mutate(match_stage = "cert_original")

# 2. 新一轮 2526 certificate 成功匹配的部分
cert_success_2526 <- cert2526_final_match %>%
  filter(!is.na(CERTIFICATE_2526)) %>%
  mutate(match_stage = "cert_2526")
```

```{r}
cert_final_match_updated <- bind_rows(
  cert_success_old,
  cert_success_2526
)
```

```{r}
cert_final_no_match_updated <- cert2526_final_match %>%
  filter(is.na(CERTIFICATE_2526)) %>%
  select(names(cert2526_final_no_match))
```

## this section: final_no_match I'll use the email variable  to perform the merge again.



```{r}
# Step 1: 先清理 final_no_match 里的 5 个 email 变量
df_email_no_match <- cert_final_no_match_updated %>%
  mutate(
    literacy_email_row_id = row_number(),
    
    username_clean = str_to_lower(str_trim(as.character(username))),
    username_2_clean = str_to_lower(str_trim(as.character(username_2))),
    WORKEMAIL_1_clean = str_to_lower(str_trim(as.character(WORKEMAIL_1))),
    WORKEMAIL_2_clean = str_to_lower(str_trim(as.character(WORKEMAIL_2))),
    WORKEMAIL_3_clean = str_to_lower(str_trim(as.character(WORKEMAIL_3)))
  )
```

```{r}
# Step 2: 清理 df_cert 里的 EMAIL_ADDRESS_2526
df_cert_email <- staff_final_merged %>%
  mutate(
    EMAIL_ADDRESS_2526_clean = str_to_lower(
      str_trim(as.character(EMAIL_ADDRESS_2526))
    )
  ) %>%
  filter(
    !is.na(EMAIL_ADDRESS_2526_clean),
    EMAIL_ADDRESS_2526_clean != ""
  ) %>%
  distinct(EMAIL_ADDRESS_2526_clean, .keep_all = TRUE)
```

```{r}
email_match_1 <- df_email_no_match %>%
  left_join(
    df_cert_email,
    by = c("username_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_username",
      NA_character_
    )
  )

email_no_match_1 <- email_match_1 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_2 <- email_no_match_1 %>%
  left_join(
    df_cert_email,
    by = c("username_2_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_username_2",
      NA_character_
    )
  )

email_no_match_2 <- email_match_2 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_3 <- email_no_match_2 %>%
  left_join(
    df_cert_email,
    by = c("WORKEMAIL_1_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_WORKEMAIL_1",
      NA_character_
    )
  )

email_no_match_3 <- email_match_3 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_4 <- email_no_match_3 %>%
  left_join(
    df_cert_email,
    by = c("WORKEMAIL_2_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_WORKEMAIL_2",
      NA_character_
    )
  )

email_no_match_4 <- email_match_4 %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_match_5 <- email_no_match_4 %>%
  left_join(
    df_cert_email,
    by = c("WORKEMAIL_3_clean" = "EMAIL_ADDRESS_2526_clean")
  ) %>%
  mutate(
    match_type = if_else(
      !is.na(EMAIL_ADDRESS_2526),
      "matched_by_WORKEMAIL_3",
      NA_character_
    )
  )
```

```{r}
email_final_match <- bind_rows(
  email_match_1 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_2 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_3 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_4 %>% filter(!is.na(EMAIL_ADDRESS_2526)),
  email_match_5
)
```

```{r}
email_final_no_match <- email_final_match %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>%
  select(names(df_email_no_match))
```

```{r}
email_final_match %>%
  count(match_type) %>% 
  print(n = Inf, width = Inf)

```

There are still about 1840 records here that cannot be merged. Given the situation, I think merging based solely on first and last name is too risky. I am considering whether we should include the MSID to create a new combination of first name, last name, school, and district to complete the merging.

```{r}
# 第一轮：原始 certificate match 成功
cert_success_1 <- cert_final_match %>%
  filter(!is.na(CERTIFICATE_FILE_NUMBER)) %>%
  mutate(
    match_stage = "certificate_original",
    final_match_type = match_type
  )

# 第二轮：CERTIFICATE_2526 match 成功
cert_success_2 <- cert2526_final_match %>%
  filter(!is.na(CERTIFICATE_2526)) %>%
  mutate(
    match_stage = "certificate_2526",
    final_match_type = match_type_2526
  )

# 第三轮：email match 成功
email_success <- email_final_match %>%
  filter(!is.na(EMAIL_ADDRESS_2526)) %>%
  mutate(
    match_stage = "email",
    final_match_type = match_type
  )
```

```{r}
all_success <- bind_rows(
  cert_success_1,
  cert_success_2,
  email_success
)
```

```{r}
all_failed <- email_final_match %>%
  filter(is.na(EMAIL_ADDRESS_2526)) %>% 
  select(names(df_email_no_match))
```

```{r}
all_success %>%
  count(match_stage, sort = TRUE) %>% 
  print(n = Inf, width = Inf)
```


## name+district+ school

```{r}
all_failed %>% 
  count(COUNTY_1)
```

```{r}
# Function: add suffix to all variables except key columns
# 功能：
# 给所有列加后缀，但保留 key（默认 TEACHER_ID 不变）
# =========================================================
add_suffix_except <- function(df, suffix, keep = "TEACHER_ID") {
  
  df_renamed <- df %>%
    rename_with(
      .fn = ~ ifelse(.x %in% keep, .x, paste0(.x, suffix))
    )
  
  return(df_renamed)
}
```

```{r}
msid <- msid_all_school %>% 
  add_suffix_except(
    suffix = "_msid"
  )
```

```{r}
msid_district <- msid %>%
  mutate(
    district_name_clean = str_to_upper(str_trim(DISTRICT_NAME_msid))
  ) %>%
  filter(
    !is.na(district_name_clean),
    district_name_clean != ""
  ) %>%
  distinct(
    district_name_clean,
    DISTRICT_msid,
    .keep_all = TRUE
  )

```

```{r}
msid_district <- msid_district %>% 
  select(DISTRICT_msid,district_name_clean)
```

```{r}
# Step 2: 清理 all_failed 里的 COUNTY_1/2/3
all_failed_county <- all_failed %>%
  mutate(
    failed_row_id = row_number(),
    
    COUNTY_1_clean = str_to_upper(str_trim(as.character(COUNTY_1))),
    COUNTY_2_clean = str_to_upper(str_trim(as.character(COUNTY_2))),
    COUNTY_3_clean = str_to_upper(str_trim(as.character(COUNTY_3)))
  )
```

```{r}
county_match_1 <- all_failed_county %>%
  left_join(
    msid_district,
    by = c("COUNTY_1_clean" = "district_name_clean")
  ) %>%
  mutate(
    county_match_type = if_else(
      !is.na(DISTRICT_msid),
      "matched_by_COUNTY_1",
      NA_character_
    )
  )

county_no_match_1 <- county_match_1 %>%
  filter(is.na(DISTRICT_msid)) %>%
  select(names(all_failed_county))
```

```{r}
county_match_2 <- county_no_match_1 %>%
  left_join(
    msid_district,
    by = c("COUNTY_2_clean" = "district_name_clean")
  ) %>%
  mutate(
    county_match_type = if_else(
      !is.na(DISTRICT_msid),
      "matched_by_COUNTY_2",
      NA_character_
    )
  )

county_no_match_2 <- county_match_2 %>%
  filter(is.na(DISTRICT_msid)) %>%
  select(names(all_failed_county))
```

```{r}
county_match_3 <- county_no_match_2 %>%
  left_join(
    msid_district,
    by = c("COUNTY_3_clean" = "district_name_clean")
  ) %>%
  mutate(
    county_match_type = if_else(
      !is.na(DISTRICT_msid),
      "matched_by_COUNTY_3",
      NA_character_
    )
  )
```

```{r}
all_failed_with_district <- bind_rows(
  county_match_1 %>% filter(!is.na(DISTRICT_msid)),
  county_match_2 %>% filter(!is.na(DISTRICT_msid)),
  county_match_3
)
```

```{r}
all_failed_with_district %>%
  count(county_match_type) %>%
  print(n = Inf)
```
```{r}
msid_school <- msid %>%
  filter(ACTIVITY_CODE_msid =="A") %>% 
  filter(
    !is.na(DISTRICT_msid),
    !is.na(SCHOOL_msid),
  ) %>%
  distinct(
    DISTRICT_msid,
    SCHOOL_msid,
    SCHOOL_NAME_LONG_msid,
    .keep_all = TRUE
  ) %>%
  select(
    DISTRICT_msid,
    SCHOOL_msid,
    SCHOOL_NAME_LONG_msid
   
  )
```

```{r}
all_failed_with_district_school <- all_failed_with_district %>%
  mutate(
    SCHOOL_1_clean = str_to_upper(str_trim(as.character(SCHOOL_1))),
    SCHOOL_2_clean = str_to_upper(str_trim(as.character(SCHOOL_2))),
    SCHOOL_3_clean = str_to_upper(str_trim(as.character(SCHOOL_3))),
    SCHOOL_4_clean = str_to_upper(str_trim(as.character(SCHOOL_4)))
  )

```

```{r}
school_match_1 <- all_failed_with_district_school %>%
  left_join(
    msid_school,
    by = c(
      "DISTRICT_msid" = "DISTRICT_msid",
      "SCHOOL_1_clean" = "SCHOOL_NAME_LONG_msid"
    )
  ) %>%
  mutate(
    school_match_type = if_else(
      !is.na(SCHOOL_msid),
      "matched_by_SCHOOL_1",
      NA_character_
    )
  )
```
```{r}
school_no_match_1 <- school_match_1 %>%
  filter(is.na(SCHOOL_msid)) %>%
  select(names(all_failed_with_district_school))
```
QC duplication
```{r}
x_dup_keys <- all_failed_with_district_school %>%
  count(DISTRICT_msid, SCHOOL_1_clean, name = "n_x") %>%
  filter(n_x > 1)

x_dup_keys
```


```{r}
y_dup_keys <- msid_school %>%
  count(DISTRICT_msid, SCHOOL_NAME_LONG_msid, name = "n_y") %>%
  filter(n_y > 1)

y_dup_keys
```
```{r}
many_to_many_keys <- x_dup_keys %>%
  inner_join(
    y_dup_keys,
    by = c(
      "DISTRICT_msid" = "DISTRICT_msid",
      "SCHOOL_1_clean" = "SCHOOL_NAME_LONG_msid"
    )
  ) %>%
  mutate(
    inflated_rows = n_x * n_y
  ) %>%
  arrange(desc(inflated_rows))

many_to_many_keys

```
```{r}
many_to_many_detail_x <- all_failed_with_district_school %>%
  semi_join(
    many_to_many_keys,
    by = c(
      "DISTRICT_msid",
      "SCHOOL_1_clean"
    )
  ) %>%
  arrange(DISTRICT_msid, SCHOOL_1_clean)

many_to_many_detail_y <- msid_school %>%
  semi_join(
    many_to_many_keys,
    by = c(
      "DISTRICT_msid",
      "SCHOOL_NAME_LONG_msid" = "SCHOOL_1_clean"
    )
  ) %>%
  arrange(DISTRICT_msid, SCHOOL_NAME_LONG_msid)
```

```{r}
View(many_to_many_detail_x)
View(many_to_many_detail_y)
```

59	7004	A SEMINOLE COUNTY VIRTUAL SCHOOL
59	7023	A SEMINOLE COUNTY VIRTUAL SCHOOL
caused duplication.


```{r}
school_match_2 <- school_no_match_1 %>%
  left_join(
    msid_school,
    by = c(
      "DISTRICT_msid" = "DISTRICT_msid",
      "SCHOOL_2_clean" = "SCHOOL_NAME_LONG_msid"
    )
  ) %>%
  mutate(
    school_match_type = if_else(
      !is.na(SCHOOL_msid),
      "matched_by_SCHOOL_2",
      NA_character_
    )
  )

school_no_match_2 <- school_match_2 %>%
  filter(is.na(SCHOOL_msid)) %>%
  select(names(all_failed_with_district_school))
```

```{r}

school_match_3 <- school_no_match_2 %>%
  left_join(
    msid_school,
    by = c(
      "DISTRICT_msid" = "DISTRICT_msid",
      "SCHOOL_3_clean" = "SCHOOL_NAME_LONG_msid"
    )
  ) %>%
  mutate(
    school_match_type = if_else(
      !is.na(SCHOOL_msid),
      "matched_by_SCHOOL_3",
      NA_character_
    )
  )

school_no_match_3 <- school_match_3 %>%
  filter(is.na(SCHOOL_msid)) %>%
  select(names(all_failed_with_district_school))
```

```{r}

school_match_4 <- school_no_match_3 %>%
  left_join(
    msid_school,
    by = c(
      "DISTRICT_msid" = "DISTRICT_msid",
      "SCHOOL_4_clean" = "SCHOOL_NAME_LONG_msid"
    )
  ) %>%
  mutate(
    school_match_type = if_else(
      !is.na(SCHOOL_msid),
      "matched_by_SCHOOL_4",
      NA_character_
    )
  )

school_no_match_4 <- school_match_4 %>%
  filter(is.na(SCHOOL_msid)) %>%
  select(names(all_failed_with_district_school))
```


```{r}
all_failed_with_school_id <- bind_rows(
  school_match_1 %>% filter(!is.na(SCHOOL_msid)),
  school_match_2 %>% filter(!is.na(SCHOOL_msid)),
  school_match_3 %>% filter(!is.na(SCHOOL_msid)),
  school_match_4
)
```

```{r}
all_failed_with_school_id %>%
  count(school_match_type) %>%
  print(n = Inf)
```
```{r}
all_failed_with_school_id %>%
  filter(is.na(SCHOOL_msid)) %>%
  count(COUNTY_1, SCHOOL_1, SCHOOL_2, SCHOOL_3, SCHOOL_4, sort = TRUE) %>%
  print(n = 100)
```



If we only use first name, last name, and district number for the merge, we still have 1,439 observations that failed to match. 

The new strategy uses First Name, Last Name, District Number, and School Number for this section.

```{r}
# Step 1: 清理 all_failed_with_district_school_id
failed_name_district_sch <- all_failed_with_school_id %>%
  mutate(
    NAME_key_first_clean = str_to_upper(str_trim(as.character(NAME_key_first))),
    NAME_key_last_clean  = str_to_upper(str_trim(as.character(NAME_key_last))),
    DISTRICT_msid_clean  = 
      as.character(DISTRICT_msid),
    SCHOOL_msid_clean = as.character(SCHOOL_msid)
      )
```

```{r}
# Step 2: 清理 staff_final_merged
staff_name_district <- staff_final_merged %>%
  mutate(
    FIRST_NAME_clean = str_to_upper(str_trim(as.character(FIRST_NAME))),
    LAST_NAME_clean  = str_to_upper(str_trim(as.character(LAST_NAME))),
    DISTRICT_SERVICE_2526_clean =  as.character(DISTRICT_SERVICE_2526),
    SCHOOL_SERVICE_2526_clean = as.character(SCHOOL_SERVICE_2526)
    ) %>%
  filter(
    !is.na(FIRST_NAME_clean),
    FIRST_NAME_clean != "",
    !is.na(LAST_NAME_clean),
    LAST_NAME_clean != "",
    !is.na(DISTRICT_SERVICE_2526_clean),
    DISTRICT_SERVICE_2526_clean != "",
    !is.na(SCHOOL_SERVICE_2526_clean),
    SCHOOL_SERVICE_2526_clean !=""
  )
```

```{r}
staff_name_district %>%
  count(FIRST_NAME_clean, LAST_NAME_clean, DISTRICT_SERVICE_2526_clean,SCHOOL_SERVICE_2526_clean, sort = TRUE) %>%
  filter(n > 1)
```
```{r}
name_district_sch_match <- failed_name_district_sch %>%
  left_join(
    staff_name_district,
    by = c(
      "NAME_key_first_clean" = "FIRST_NAME_clean",
      "NAME_key_last_clean"  = "LAST_NAME_clean",
      "DISTRICT_msid_clean"  = "DISTRICT_SERVICE_2526_clean",
      "SCHOOL_msid_clean" = "SCHOOL_SERVICE_2526_clean"
    )
  ) 
```

```{r}
name_district_sch_no_match <- name_district_sch_match %>%
  filter(is.na(TEACHER_ID))
```

```{r}
name_combination_success <- name_district_sch_match %>%
  filter(!is.na(TEACHER_ID)) %>%
  mutate(
    match_stage = "name_district_school",
    final_match_type = "matched_by_name_and_district_sch"
  )
```

```{r}

all_success_updated <- bind_rows(
  all_success,
  name_combination_success
)
```
```{r}
names(all_success_updated)
```

```{r}
all_failed_updated <- name_district_sch_match %>%
  filter(is.na(TEACHER_ID))
```

```{r}
all_success_updated %>% count(final_match_type)
```
```{r}
all_success_updated %>% count(
  username
) %>% filter(n>1)
```

```{r}
write.csv(all_failed_updated,
          "failed_get_teacherid.csv",
          row.names = FALSE)
```

```{r}
write.csv(all_success_updated,
          "treatment_with_teacherid.csv",
          row.names = FALSE)
```

## Section 3 combing student data

```{r}
#Load TeacherCourse locally.
teacher_course_local <- teacher_course %>% collect()
```

```{r}
teacher_course_local <- teacher_course_local %>% 
  add_suffix_except(
  suffix = "_tc",
  keep = "TEACHER_ID"
  )
```


```{r}
treatment_data_v1 <- all_success_updated %>% 
  left_join(teacher_course_local,
            by= "TEACHER_ID")
```
```{r}
all_success_updated <- all_success_updated %>%
  mutate(
    DISTRICT_SERVICE = as.character(
      as.numeric(DISTRICT_SERVICE)
    )
  )
```

```{r}
treatment_data_v2 <- all_success_updated %>% 
  left_join(teacher_course_local,
            by=c("TEACHER_ID",
            "DISTRICT_SERVICE"="DISTRICT_INST_tc"))
```

```{r}
treatment_data_v1 %>%
  count(TEACHER_ID, DISTRICT_SERVICE,COURSE_tc) %>%
  filter(n > 1)
```
```{r}
treatment_data_v2 %>%
  count(TEACHER_ID, DISTRICT_SERVICE,COURSE_tc) %>%
  filter(n > 1)
```

### Qc duplication
for v1 only teacherid

```{r}
# v1 join key: TEACHER_ID

x_dup_v1 <- all_success_updated %>%
  count(TEACHER_ID, name = "n_x") %>%
  filter(n_x > 1)

y_dup_v1 <- teacher_course_local %>%
  count(TEACHER_ID, name = "n_y") %>%
  filter(n_y > 1)

explosion_keys_v1 <- x_dup_v1 %>%
  inner_join(y_dup_v1, by = "TEACHER_ID") %>%
  mutate(exploded_rows = n_x * n_y) %>%
  arrange(desc(exploded_rows))

explosion_keys_v1
```
```{r}
explosion_x_v1 <- all_success_updated %>%
  semi_join(explosion_keys_v1, by = "TEACHER_ID") %>%
  arrange(TEACHER_ID)

View(explosion_x_v1)
```

6U2FN3YL0 trouble shooting in original data. same cert_1 1401640

```{r}
# v2 join key: TEACHER_ID + DISTRICT

x_dup_v2 <- all_success_updated %>%
  count(TEACHER_ID, DISTRICT_SERVICE, name = "n_x") %>%
  filter(n_x > 1)

y_dup_v2 <- teacher_course_local %>%
  count(TEACHER_ID, DISTRICT_INST_tc, name = "n_y") %>%
  filter(n_y > 1)

explosion_keys_v2 <- x_dup_v2 %>%
  inner_join(
    y_dup_v2,
    by = c(
      "TEACHER_ID" = "TEACHER_ID",
      "DISTRICT_SERVICE" = "DISTRICT_INST_tc"
    )
  ) %>%
  mutate(exploded_rows = n_x * n_y) %>%
  arrange(desc(exploded_rows))

explosion_keys_v2
```
same issue. 

```{r}
treatment_data_v1 %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  summarise(n = n())
```

```{r}
treatment_no_studentinfo <- treatment_data_v1 %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc))
```

```{r}
missing_keys_v1 <- treatment_data_v1 %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  select(TEACHER_ID, DISTRICT_SERVICE) %>%
  distinct()
```
```{r}
missing_keys_v1 <- missing_keys_v1 %>% 
  mutate(DISTRICT_SERVICE= as.character(DISTRICT_SERVICE))
```

double check for v2

```{r}
missing_keys_v2 <- treatment_data_v2 %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  select(TEACHER_ID, DISTRICT_SERVICE) %>%
  distinct()
```
```{r}
missing_keys_v2 <- missing_keys_v2 %>% 
  mutate(DISTRICT_SERVICE= as.character(DISTRICT_SERVICE))
```


```{r}
teacher_course_check_v2<- teacher_course_local %>%
  semi_join(
    missing_keys_v2,
    by = c(
      "TEACHER_ID",
      "DISTRICT_INST_tc" = "DISTRICT_SERVICE"
    )
  )
```

```{r}
teacher_course_wrong_district <- teacher_course_local %>%
  semi_join(
    missing_keys_v2,
    by = "TEACHER_ID"
  ) %>%
  select(TEACHER_ID, DISTRICT_INST_tc, SCHOOL_INST_tc, COURSE_tc, FLEID_tc) %>%
  distinct()
```

```{r}
missing_keys %>%
  left_join(
    teacher_course_wrong_district,
    by = "TEACHER_ID"
  )
```

```{r}
write.csv(treatment_no_studentinfo,
        "teachers_without_students_info.csv",
       row.names = FALSE)
```

The inspection is complete. I have confirmed that these teacher IDs are not in the teacher Courses data. need to double check and discuss how to handle this issue.

After completing the merge, we discovered another issue: the district service and district instruction fields are inconsistent.

How should we handle this? We need to reconcile these differences.

next , I will combine student demo and deduplication

```{r}
treatment_data <- treatment_data_v1%>% 
  mutate(
    DISTRICT_tc = as.numeric(as.character(DISTRICT_tc)),
    SCHOOL_tc   = as.numeric(as.character(SCHOOL_tc)),
    SchoolID_tc = DISTRICT_tc * 10000 + SCHOOL_tc,
    UniqueID    = paste0(FLEID_tc, SchoolID_tc)
  )
```

```{r}
student_demo_local <- student_demo %>% collect()
```

```{r}
student_demo_local <- student_demo_local %>% 
  add_suffix_except(
  suffix = "_demo",
  keep = "FLEID"
  )
```

```{r}
student_demo_local <- student_demo_local %>% 
  mutate(
    DISTRICT_demo = as.numeric(as.character(DISTRICT_demo)),
    SCHOOL_demo   = as.numeric(as.character(SCHOOL_demo)),
    SchoolID_demo = DISTRICT_demo * 10000 + SCHOOL_demo,
    UniqueID    = paste0(FLEID, SchoolID_demo)
  )
```

```{r}
dup_in_demo <- student_demo_local %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  pull(UniqueID)
```

deduplication strategy 1 :deduplication in student_demo drop district_inst ==71

```{r}
student_demo_clean <- student_demo_local %>%
  group_by(UniqueID) %>%
  filter(!(n() > 1 & DISTRICT_INST_demo == 71)) %>%
  ungroup()
```

```{r}
nrow(student_demo_local)
nrow(student_demo_clean)
```

```{r}
student_demo_clean %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  arrange(desc(n))# need to manually check

```
```{r}
student_demo_clean %>% 
  filter(UniqueID=="FL000000126868137091")
```

```{r}
treatment_data_final <- treatment_data %>% 
  left_join(student_demo_clean,
            by="UniqueID")
```
```{r}
# 1. 左表 treatment_data 里重复的 UniqueID
x_dup_uid <- treatment_data %>%
  count(UniqueID, name = "n_x") %>%
  filter(!is.na(UniqueID), n_x > 1)

# 2. 右表 student_demo_clean 里重复的 UniqueID
y_dup_uid <- student_demo_clean %>%
  count(UniqueID, name = "n_y") %>%
  filter(!is.na(UniqueID), n_y > 1)

# 3. 真正造成 many-to-many / data explosion 的 UniqueID
explosion_uid <- x_dup_uid %>%
  inner_join(
    y_dup_uid,
    by = "UniqueID"
  ) %>%
  mutate(
    exploded_rows = n_x * n_y
  ) %>%
  arrange(desc(exploded_rows))

explosion_uid
```
```{r}
treatment_data %>% 
  filter(UniqueID=="FL000000774074290089")

```
```{r}
student_demo_clean %>% 
 filter(UniqueID=="FL000000774074290089")
```


only 7 data explosion

```{r}
 treatment_data_final %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  arrange(desc(n)) %>% 
   print(Inf)
```

```{r}
treatment_data_final %>% 
  filter(UniqueID =="FL000007593932180301")
```

```{r}
write.csv(treatment_data_final,
          "main_treatment_data_v3.csv",
          row.names = FALSE)
```

```{r}
dbDisconnect(con)
```
