---
title: "25_26_doe_data"
---

## Import package and trying to connect snowflake

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(purrr)
library(stringdist)
library(openxlsx)
library(stringr)
#The following package is for connecting to Snowflake.
library(DBI)
library(odbc)
library(openssl)  # To read private key
```

## Connect to Snowflake

```{r}
# Connect to Snowflake using key-pair authentication

con <- dbConnect(odbc::odbc(),
                 Driver = "SnowflakeDSIIDriver", #from UFIT
                 Server = "ufl-datahub.snowflakecomputing.com", #from LC Data team
                 UID = "xu.rui@ufl.edu", #from UFIT
                 authenticator = "externalbrowser", #from UFIT
                 Warehouse     = "REPORTING", #from LC Data team
                 Database      = "PUB_LAKE_LASTINGER" #can specify specific database needing to access
                #Schema        = "ADD" #can specify specific Schema - aka folders - needing to access
)

# Look in Connections - top right RStudio - for Database/Schema list
```

## Import data in snowflake

```{r}
stff <- tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_STAFF"))
```

```{r}
staff_suffix <- tbl(con,dbplyr::in_schema("RAW_DOE","02_13_2026_PERA_4472B_STAFF_CERTIFICATION_DATA_2025_26_FS2"))
```

```{r}
Literacy_matrix <-  tbl(con,dbplyr::in_schema("LAKE_LASTINGER","SP_25_26_STIPENDS_LITERACY_MATRIX_FY25_26"))
```

```{r}
teacher_course <- tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_TEACHER_COURSE"))
```

```{r}
student_demo <-  tbl(con,dbplyr::in_schema("RAW_DOE","01_15_2026_STUDENT_DEMO"))
```

## explore

```{r}
stff %>%
  count(TEACHER_ID) %>%
  filter(n > 1)
```

```{r}
qc_teacher_id <- stff %>%
  summarise(
    total_rows = n(),
    unique_teacher = n_distinct(TEACHER_ID),
    missing_teacher = sum(ifelse(is.na(TEACHER_ID), 1L, 0L), na.rm = TRUE)#note: The is.na function cannot be recognized in Snowflake. Therefore, we must switch to a method that is compatible with sql.
  ) %>%
  collect()

qc_teacher_id


```

```{r}
qc_teacher_id_1 <- staff_suffix %>%
  summarise(
    total_rows = n(),
    unique_teacher = n_distinct(TEACHER_ID),
    missing_teacher = sum(ifelse(is.na(TEACHER_ID), 1L, 0L), na.rm = TRUE)#same issue
  ) %>%
  collect()

qc_teacher_id_1

```

Filter out duplicate data\
function:

```{r}


# =========================================================
# Function: split unique TEACHER_ID rows and duplicated rows
# 功能：
# 1. 根据 TEACHER_ID 统计重复情况
# 2. 保留唯一 TEACHER_ID 的数据
# 3. 把重复 TEACHER_ID 的数据单独保存
# 4. 给数据加上来源标签 source
# =========================================================
split_teacher_data <- function(df, source_name) {
  
  # 第一步：计算每个 TEACHER_ID 出现的次数
  teacher_count <- df %>%
    count(TEACHER_ID, name = "teacher_id_n")
  
  # 第二步：把次数合并回原始数据
  df_with_count <- df %>%
    left_join(teacher_count, by = "TEACHER_ID")
  
  # 第三步：保留 TEACHER_ID 只出现 1 次的行
  df_unique <- df_with_count %>%
    filter(teacher_id_n == 1) %>%
    mutate(source = source_name)
  
  # 第四步：把重复的 TEACHER_ID 行单独保存
  df_duplicate <- df_with_count %>%
    filter(teacher_id_n > 1) %>%
    mutate(source = source_name)
  
  # 第五步：重复 TEACHER_ID 的 summary（每个ID重复几次）
  dup_summary <- teacher_count %>%
    filter(teacher_id_n > 1) %>%
    arrange(desc(teacher_id_n))
  
  # 返回一个 list，方便后续调用
  return(list(
    unique_data = df_unique,
    duplicate_data = df_duplicate,
    duplicate_summary = dup_summary
  ))
}
```

```{r}
# 对第一个数据库处理
staff_2526_result <- split_teacher_data(
  df = stff,
  source_name = "stff"
)

# 对第二个数据库处理
staff_july_result <- split_teacher_data(
  df = staff_suffix,
  source_name = "staff_suffix"
)
```

```{r}
# 第一个数据库：唯一 / 重复 / 重复summary
staff_2526_unique <- staff_2526_result$unique_data
staff_2526_duplicate <- staff_2526_result$duplicate_data
staff_2526_dup_summary <- staff_2526_result$duplicate_summary

# 第二个数据库：唯一 / 重复 / 重复summary
staff_july_unique <- staff_july_result$unique_data
staff_july_duplicate <- staff_july_result$duplicate_data
staff_july_dup_summary <- staff_july_result$duplicate_summary
```

export duplication

```{r}
write.csv(staff_2526_duplicate, "staff_2526_duplicate.csv", row.names = FALSE)
write.csv(staff_july_duplicate, "staff_suffix.csv", row.names = FALSE)
```

export unique teacher\

```{r}
write.csv(staff_2526_unique,"staff_2526_uqniue.csv", row.names = FALSE)
write.csv(staff_july_unique,"staff_suffix_unique.csv", row.names = FALSE)
```

function add suffix

```{r}
# Function: add suffix to all variables except key columns
# 功能：
# 给所有列加后缀，但保留 key（默认 TEACHER_ID 不变）
# =========================================================
add_suffix_except <- function(df, suffix, keep = "TEACHER_ID") {
  
  df_renamed <- df %>%
    rename_with(
      .fn = ~ ifelse(.x %in% keep, .x, paste0(.x, suffix))
    )
  
  return(df_renamed)
}
```

```{r}
staff_2526_unique_sfx <- add_suffix_except(
  df = staff_2526_unique,
  suffix = "_2526"
)
```

## unique data merge

```{r}
staff_merged <- staff_july_unique %>%
  full_join(
    staff_2526_unique_sfx,
    by = "TEACHER_ID"
  )
```

## qc

```{r}
staff_merged %>%
  summarise(
    total_rows = n(),
    unique_teacher = n_distinct(TEACHER_ID)
  )
```

```{r}

staff_merged_qc <- staff_merged %>%
  mutate(
    source_flag = case_when(
      !is.na(source) & !is.na(source_2526) ~ "both",
      !is.na(source) & is.na(source_2526) ~ "only_july",
      is.na(source) & !is.na(source_2526) ~ "only_2526",
      TRUE ~ NA_character_
    )
  )

```

## For duplication data merge

Check the combination of district and teacher id

```{r}
staff_july_duplicate_check <- staff_july_duplicate %>%
  count(TEACHER_ID, DISTRICT_SERVICE, name = "n") %>%
  filter(n > 1) %>%
  arrange(desc(n))

staff_july_duplicate_check
```

```{r}
staff_2526_duplicate_check <- staff_2526_duplicate %>%
  count(TEACHER_ID, DISTRICT_SERVICE, name = "n") %>%
  filter(n > 1) %>%
  arrange(desc(n))

staff_2526_duplicate_check
```

```{r}
staff_july_duplicate %>%
  summarise(
    total_rows = n(),
    distinct_teacher = n_distinct(TEACHER_ID),
    distinct_teacher_district = n_distinct(paste(TEACHER_ID, DISTRICT_SERVICE, sep = "__"))
  )
```

```{r}
staff_2526_duplicate %>%
  summarise(
    total_rows = n(),
    distinct_teacher = n_distinct(TEACHER_ID),
    distinct_teacher_district = n_distinct(paste(TEACHER_ID, DISTRICT_SERVICE, sep = "__"))
  )
```

```{r}
nrow(staff_july_duplicate_check)
nrow(staff_2526_duplicate_check)
```

```{r}
staff_2526_duplicate_sfx <- add_suffix_except(
  df = staff_2526_duplicate,
  suffix = "_2526",
  keep = c("TEACHER_ID", "DISTRICT_SERVICE")
)
```

```{r}
staff_duplicate_merged <- staff_july_duplicate %>%
  full_join(
    staff_2526_duplicate_sfx,
    by = c("TEACHER_ID", "DISTRICT_SERVICE")
  )
```

```{r}
staff_duplicate_merged <- staff_duplicate_merged %>%
  mutate(
    source_flag = case_when(
      !is.na(source) & !is.na(source_2526) ~ "both",
      !is.na(source) & is.na(source_2526) ~ "only_july",
      is.na(source) & !is.na(source_2526) ~ "only_2526",
      TRUE ~ NA_character_
    )
  )
```

```{r}
staff_duplicate_merged %>%
  count(source_flag)
```

```{r}
setdiff(colnames(staff_merged), colnames(staff_duplicate_merged))
setdiff(colnames(staff_duplicate_merged), colnames(staff_merged))
```

```{r}
staff_duplicate_merged <- staff_duplicate_merged %>%
  mutate(DISTRICT_SERVICE_2526 = DISTRICT_SERVICE)
```

```{r}
#note: The bind_row function cannot be used directly on Snowflake. Therefore, we need to replace it with a compatible Snowflake function called union_all.
staff_final_merged <- union_all(
  staff_merged,
  staff_duplicate_merged
) %>% 
  collect()
```

```{r}
staff_final_merged %>% count(teacher_id_n_2526)
```

```{r}
write.csv(staff_final_merged,
      "survey2_2526_twodatasets_merge.csv",
row.names = FALSE)
```

The section above successfully merged the two staff databases, and the results look good.

## Next, I will use the literacy matrix to merge the staff datasets.

```{r}
df_dcf <- Literacy_matrix %>%
  mutate(
    DCF_clean = str_replace_all(as.character(DCF), "[^0-9]", "")
  ) %>% collect()

df_cert <- staff_final_merged %>%
  mutate(
    Cert_clean = str_replace_all(as.character(CERTIFICATE_FILE_NUMBER), "[^0-9]", "")
  )

```

Exact match

```{r}
exact_match <- df_dcf %>%
  left_join(
    df_cert,
    by = c("DCF_clean" = "Cert_clean")
  ) %>%
  mutate(match_type = if_else(!is.na(CERTIFICATE_FILE_NUMBER), "exact_match", NA_character_))
```

Find the failed match

```{r}
dcf_no_match <- exact_match %>%
  filter(is.na(CERTIFICATE_FILE_NUMBER)) %>%
  select(names(df_dcf))
```

```{r}
write.csv(exact_match,"Literacy_Matrix_with_staff_inf.csv",row.names = FALSE)
```

```{r}
write.csv(dcf_no_match,"failed_mathch_by_DCF.csv",row.names = FALSE)
```

I have manually checked it, and the field match looks acceptable.

The next step will be to merge the teacher courses and student demographics.

note:

Currently, the workflow is still a mix of Snowflake and R.

My preliminary plan is to handle the data merging in Snowflake, as it would be much more efficient. However, since the code hasn't been converted yet, I will first validate my logic and generate the results before finding a way to refactor the code.

```{r}
#Load TeacherCourse locally.
teacher_course_local <- teacher_course %>% collect()

```

```{r}
teacher_course_local <- teacher_course_local %>% 
  add_suffix_except(
  suffix = "_tc",
  keep = "TEACHER_ID"
  )
```

```{r}
treatment_data <- exact_match %>% 
  left_join(teacher_course_local,
            by=c("TEACHER_ID",
            "DISTRICT_SERVICE"="DISTRICT_INST_tc"))
```

```{r}
treatment_data %>%
  count(TEACHER_ID, DISTRICT_SERVICE) %>%
  filter(n > 1)
```

Check the database here for instances where a teacherID exists, but student information is missing.

```{r}
treatment_data %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  summarise(n = n())
```

```{r}
treatment_no_studentinfo <- treatment_data %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc))
```

```{r}
missing_fle <- treatment_data %>%
  filter(!is.na(TEACHER_ID) & is.na(FLEID_tc)) %>%
  select(TEACHER_ID) %>%
  distinct()
```

```{r}
teacher_course_check <- teacher_course_local %>%
  filter(TEACHER_ID %in% missing_fle$TEACHER_ID)
```

manually check

```{r}
teacher_course_local %>%
  filter(TEACHER_ID == "OX7752QWM") %>% 
  nrow()
```

The inspection is complete. I have confirmed that these teacher IDs are not in the teacher Courses data. need to double check and discuss how to handle this issue.

### in this section I will merge student demo and duplication

```{r}
names(treatment_data)
```

```{r}
treatment_data <- treatment_data %>% 
  mutate(
    DISTRICT_tc = as.numeric(as.character(DISTRICT_tc)),
    SCHOOL_tc   = as.numeric(as.character(SCHOOL_tc)),
    SchoolID_tc = DISTRICT_tc * 10000 + SCHOOL_tc,
    UniqueID    = paste0(FLEID_tc, SchoolID_tc)
  )
```

```{r}
student_demo_local <- student_demo %>% collect()
```

```{r}
student_demo_local <- student_demo_local %>% 
  add_suffix_except(
  suffix = "_demo",
  keep = "FLEID"
  )
```

```{r}
student_demo_local <- student_demo_local %>% 
  mutate(
    DISTRICT_demo = as.numeric(as.character(DISTRICT_demo)),
    SCHOOL_demo   = as.numeric(as.character(SCHOOL_demo)),
    SchoolID_demo = DISTRICT_demo * 10000 + SCHOOL_demo,
    UniqueID    = paste0(FLEID, SchoolID_demo)
  )
```

```{r}
dup_in_demo <- student_demo_local %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  pull(UniqueID)
```

```{r}
student_demo_local %>%
  filter(UniqueID %in% dup_in_demo) %>%
  summarise(n_71 = sum(DISTRICT_INST_demo == 71, na.rm = TRUE))

```

deduplication in student_demo drop district_inst ==71

```{r}
student_demo_clean <- student_demo_local %>%
  group_by(UniqueID) %>%
  filter(!(n() > 1 & DISTRICT_INST_demo == 71)) %>%
  ungroup()
```

```{r}
nrow(student_demo_local)
nrow(student_demo_clean)
```

```{r}
student_demo_clean %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  arrange(desc(n))# need to manually check
```

```{r}
treatment_data_final <- treatment_data %>% 
  left_join(student_demo_clean,
            by="UniqueID")
```

```{r}
 treatment_data_final %>%
  count(UniqueID) %>%
  filter(n > 1) %>%
  arrange(desc(n))
```

```{r}
write.csv(treatment_data_final,
          "main_treatment_data.csv",
          row.names = FALSE)
```

```{r}
dbDisconnect(con)
```
